### 2023-05-09 bachelorthesis/10-motivation.tex
#39: Introduction -> Motivation
#39: Theory finished and closure corrected, First Draft finished

\chapter{Introduction}
\label{chap:intro}

Ion temperature gradient driven turbulence close to marginal stability exhibits zonal flow pattern formation on mesoscales, so-called $\exb$ staircase structures \cite{pradalier2010}.
Such pattern formation has been observed in local gradient-driven flux-tube simulations \cite{peeters2016, weikl2017, rath2021} as well as global gradient-driven \cite{mcmillan2009, villard2013, seo2022} and global flux-driven \cite{pradalier2010, pradalier2015, wang2020, kim2022, kishimoto2023} studies. 
In global studies, spanning a larger fraction of the minor radius, multiple radial repetitions of staircase structures are usually observed, with a typical pattern size of several ten Larmor radii.
% By contrast, the radial domain size of local flux-tube simulations is often restricted to a similar mesoscale.
By contrast, in the aforementioned local studies the radial size of $\exb$ staircase structures is always found to converge to the radial box size of the flux tube domain.
The above observations lead to the question: \bigskip

@@

The latter case would imply that it is not necessarily global physics, i.e., profile effects, that set 
\begin{itemize}
	\item[(i)] the radial size of the $\exb$ staircase pattern
	\item[(ii)] the scale of avalanche-like transport events.
\end{itemize}
These transport events are usually restricted to $\exb$ staircase structures and considered as a nonlocal transport mechanism \cite{pradalier2010}. \bigskip

In this bachlor thesis the above question is addressed through a box size convergence scan of the same cases close to the nonlinear threshold for turbulence generation as studied in Ref. \citenum{peeters2016}.
%TODO More Text

>>>>

\chapter{Motivation}
\label{chap:motivation}

\thispagestyle{empty}
\newpage

Ion temperature gradient driven turbulence close to marginal stability exhibits zonal flow pattern formation on mesoscales, so-called $\exb$ staircase structures \cite{Pradalier2010}.
Such pattern formation has been observed in local gradient-driven flux-tube simulations \cite{Peeters2016, Weikl2017, Rath2021} as well as global gradient-driven \cite{McMillan2009, Villard2013, Seo2022} and global flux-driven \cite{Pradalier2010, Pradalier2015, Wang2020, Kim2022, Kishimoto2023} studies. 
In global studies, spanning a larger fraction of the minor radius, multiple radial repetitions of staircase structures are usually observed, with a typical pattern size of several ten Larmor radii.
% By contrast, the radial domain size of local flux-tube simulations is often restricted to a similar mesoscale.
By contrast, in the aforementioned local studies the radial size of $\exb$ staircase structures is always found to converge to the radial box size of the flux tube domain.
The above observations lead to the question: \bigskip

@@

The latter case would imply that it is not necessarily global physics, i.e., profile effects, that set 
\begin{itemize}
	\item[(i)] the radial size of the $\exb$ staircase pattern
	\item[(ii)] the scale of avalanche-like transport events.
\end{itemize}
These transport events are usually restricted to $\exb$ staircase structures and considered as a nonlocal transport mechanism \cite{Pradalier2010}. \bigskip

In this bachlor thesis the above question is addressed through a box size convergence scan of the same cases close to the nonlinear threshold for turbulence generation as studied in Ref. \citenum{Peeters2016}.
%TODO More Text

### 2023-05-09 bachelorthesis/21-motion.tex
#39: Theory finished and closure corrected, First Draft finished

In magnetic confinement devices like the tokamak reactor, the charged particles experience magnetic and electric fields which results in distinct motion under tha associated force. Charged particles can be separated in species, e.g. electrons and ions, which will be later on indicated bei the index sp on the quantities.

\subsection{Particle Motion perpendicular to the magnetic field - The Gyro-Motion}
\label{sub:gyromotion}

Due to the Lorentz force particles with a velocity component perpendicular to the homogenous magnetic field $v_{\perp,\mathrm{sp}}$ undergo a circular motion in the plane perpendicular to the magnetic field. This type motion has circular frequency which is often referred as \textit{cyclotron frequency} and is defined as:
\begin{gather}
    \omega_\mathrm{c,sp} = \frac{|q_\mathrm{sp}|B}{m_\mathrm{sp}}~,
\end{gather}
where $m_\mathrm{sp}$ and $q_\mathrm{sp}$ are the mass and the charge of the particle and $B$ the strength of the magnetic field. The radius, the so called \textit{Larmor radius}, of this motion is given by:
\begin{gather}
    \rho_\mathrm{sp} = \frac{m_\mathrm{sp}v_{\perp,\mathrm{sp}}}{|q_\mathrm{sp}|B}
\end{gather}
which center is often as \textit{gyro-center}. Note that since the Lorentz force depends on the species of the particle the circulation direction is opposed between electron in ions.

Due to the Coulomb collisions the plasma gets thermalized. Together with the Maxwell-Boltzmann distribution the typical species thermal velocity is
\begin{gather}
    v_\mathrm{th,sp} = \sqrt{\frac{2T_\mathrm{sp}}{m_\mathrm{sp}}}~,
\end{gather}
where $T_\mathrm{sp}$ represents the species temperature. Based on the thermal velocity $v_\mathrm{th,sp}$ the \textit{thermal Larmor radius} gets introduced as:
\begin{gather}
    \rho_\mathrm{th,sp} = \frac{m_\mathrm{sp}v_{\mathrm{th,sp}}}{|q_\mathrm{sp}|B}~.
\end{gather}

@@

\subsection{Particle Motion parallel to the magnetic field - The Gyro Motion}
\label{sub:parallelmotion}

In absence of forces in the parallel direction of the magnetic field the particles can stream freely parallel to the homogenous magnetic field. The velocity of this motion is connected to the thermal velocity $v_\mathrm{th,sp}$ and is dominated bei electrons due to their lighter mass compared to ions ($v_\mathrm{th,e}/v_\mathrm{th,i} = 60$).

When an electric field with a component parallel to the magnetic field $E_\parallel$ acts on plasma the charged particles gets accelerated by the electric force
\begin{gather}
    F_{\parallel,E} = q_\mathrm{sp}E_\parallel~.
\end{gather}
The parallel motion follows then from the equation of motion. Here the direction of the motion also depends on the species type.

Since magnetic fields are not always homogenous an inhomogeneous magnetic field with its gradient $\nabla B$ contains an component parallel to the magnetic field which is given by:
\begin{gather}
    \nabla_\parallel B = \frac{\vect{B}}{B} \cdot \nabla B~.
\end{gather}
This causes the force
\begin{gather}
    F_{\parallel,\nabla_\parallel B} = - \frac{m_\mathrm{sp}v^2_{\perp,\mathrm{sp}}}{2B} \nabla_\parallel B = \mu \nabla_\parallel B~;~~~~\mu = \frac{m_\mathrm{sp}v^2_{\perp,\mathrm{sp}}}{2B}
\end{gather}
with \textit{magnetic moment} $\mu$ which can be an adabatic invariant (constant of motion) if the variation of the magnetic field over time is smaller than the inverse of the cyclotron freqency $\omega^{-1}_\mathrm{c,sp}$ and the spatial variation is larger the Larmor radius $\rho_\mathrm{sp}$. The caused force has it application in the mirror effect where a charged particle gets reflected due to this force.

@@

\subsection{Drifts in the Gyro-Center}
\label{sub:drift}

In the presence of a magnetic field (homogenous, inhomogeneous or perturbed) and electric fields the gyro-center undergoes slow (compared to the thermal velocity $\vth$) drift motions perpendicular to the magnetic field. There are serval examples for this drift motion. According to this thesis topic only the main three drift types will be covered in the following.

\begin{enumerate}
    \item \textbf{$\exb$ drift:}\\
    If an electric field $\vect{E}$ with an perpendicular component is present together with a the magnetic field $\vect{B}$ (both fields are homogenous) the acting Coulomb force and Lorentz force results to a drift of the gyro-center with
    \begin{gather}
        \vect{v}_{E} = \frac{\vect{E}\times\vect{B}}{B^2}
        \label{eq:exb}
    \end{gather}
    which is called the $\exb$-drift. Since both acting forces direction depends on the species type the direction of the $\exb$-drift is for every species the same.\\
    \item \textbf{$\nabla B$ Drift:}\\
    Inhomogeneous magnetic field causes an gradient $\nabla B$ of the magnetic field. Because of that gradient the gyro-center undergoes an $\nabla B$-drift defined by
    \begin{gather}
        \vect{v}_{\nabla B} = \frac{m_\mathrm{sp} v^2_{\perp,\mathrm{sp}}}{2 q_\mathrm{sp}}\frac{\vect{B}\times \nabla B}{B^3}~.
        \label{eq:gradB}
    \end{gather}
    The $\nabla B$-drift varies thereby on scales which are larger compared to the Larmor radius. The direction of the $\nabla B$-drift depends on the species type. \\
    \item \textbf{Curvature drift:}\\
    Due to centrifugal force acting on the particle in a curved magnetic filed the gyro-center experience an curvature drift given by
    \begin{gather}
        \vect{v}_{\kappa} = \frac{m_\mathrm{sp} v^2_{\parallel,\mathrm{sp}}}{q_\mathrm{sp}}\frac{\vect{B}\times \kappa}{B^2} = \frac{m_\mathrm{sp} v^2_{\parallel,\mathrm{sp}}}{q_\mathrm{sp}} \frac{\vect{B}\times \nabla B}{B^3}~;~~~~\kappa = \frac{\nabla B}{B}~.
        \label{eq:curvature}
    \end{gather}
    To obtain the result for $\kappa$ in Eq. \ref{eq:curvature} the plasma pressure has to be small in comparison die the magnetic field strength $B$. In the form of Eq. \ref{eq:curvature} $\nabla B$- and curvature drift can be treated similarly.
\end{enumerate}

>>>>

In magnetic confinement devices like the tokamak reactor, the charged particles experience magnetic and electric fields which results in distinct motion under tha associated force. Charged particles can be separated in species, e.g. electrons and ions, which will be later on not displayed in the governing equation. Throughout this thesis the charge $q$ or the mass $m$ indicate the quantities of a specific spiecies.

\subsection{Particle Motion perpendicular to the magnetic field}
\label{sub:gyromotion}

Due to the Lorentz force particles with a velocity component perpendicular to the homogenous magnetic field $v_{\perp}$ undergo a circular motion in the plane perpendicular to the magnetic field. This type motion has circular frequency which is often referred as \textit{cyclotron frequency} and is defined as:
\begin{gather}
    \omega_\mathrm{c} = \frac{|q|B}{m}~,
\end{gather}
where $m$ and $q$ are the mass and the charge of the particle and $B$ the strength of the magnetic field. The radius, the so called \textit{Larmor radius}, of this motion is given by:
\begin{gather}
    \rho_\mathrm{L} = \frac{mv_{\perp}}{|q|B}
\end{gather}
which center is often as \textit{gyrocenter}. Note that since the Lorentz force depends on the species of the particle the circulation direction is opposed between electron in ions.

Due to the Coulomb collisions the plasma gets thermalized. Together with the Maxwell-Boltzmann distribution the typical species thermal velocity is
\begin{gather}
    \vth = \sqrt{\frac{2T}{m}}~,
\end{gather}
where $T$ represents the species temperature. Based on the thermal velocity $v_\mathrm{th}$ the \textit{thermal Larmor radius} gets introduced as \cite{Wesson2011}
\begin{gather}
    \rhoth = \frac{m\vth}{|q|B}~.
\end{gather}

@@

\subsection{Particle Motion parallel to the magnetic field}
\label{sub:parallelmotion}

In absence of forces in the parallel direction of the magnetic field the particles can stream freely parallel to the homogenous magnetic field. The velocity of this motion is connected to the thermal velocity $\vth$ and is dominated bei electrons due to their lighter mass compared to ions ($v_\mathrm{th,e}/v_\mathrm{th,i} = 60$).

When an electric field with a component parallel to the magnetic field $E_\parallel$ acts on plasma the charged particles gets accelerated by the electric force
\begin{gather}
    F_{\parallel,E} = qE_\parallel~.
\end{gather}
The parallel motion follows then from the equation of motion. Here the direction of the motion also depends on the species type.

Since magnetic fields are not always homogenous an inhomogeneous magnetic field with its gradient $\nabla B$ contains a component parallel to the magnetic field which is given by:
\begin{gather}
    \nabla_\parallel B = \frac{\vect{B}}{B} \cdot \nabla B~.
\end{gather}
This causes the force
\begin{gather}
    F_{\parallel,\nabla_\parallel B} = - \frac{mv^2_{\perp}}{2B} \nabla_\parallel B = \mu \nabla_\parallel B~;~~~~\mu = \frac{mv^2_{\perp}}{2B}
\end{gather}
with \textit{magnetic moment} $\mu$ which can be an adabatic invariant (constant of motion) if the variation of the magnetic field over time is smaller than the inverse of the cyclotron freqency $\omega^{-1}_\mathrm{c}$ and the spatial variation is larger the Larmor radius $\rho_\mathrm{L}$. The caused force has it application in the mirror effect where a charged particle gets reflected due to this force. \cite{Wesson2011}

@@

\subsection{Drifts in the Gyrocenter}
\label{sub:drift}

In the presence of a magnetic field (homogenous, inhomogeneous or perturbed) and electric fields the gyrocenter undergoes slow (compared to the thermal velocity $\vth$) drift motions perpendicular to the magnetic field. There are serval examples for this drift motion. According to this thesis topic only the main three drift types will be covered in the following.

\begin{enumerate}
    \item \textbf{$\exb$ Drift:}\\
    If an electric field $\vect{E}$ with a perpendicular component is present together with the magnetic field $\vect{B}$ (both fields are homogenous) the acting Coulomb force and Lorentz force results to a drift of the gyrocenter with
    \begin{gather}
        \vect{v}_{E} = \frac{\vect{E}\times\vect{B}}{B^2}
        \label{eq:exb}
    \end{gather}
    which is called the $\exb$-drift. Since both acting forces direction depends on the species type the direction of the $\exb$ drift is for every species the same.
    \item \textbf{$\nabla B$ Drift:}\\
    Inhomogeneous magnetic field causes a gradient $\nabla B$ of the magnetic field. Because of that gradient the gyrocenter undergoes a $\nabla B$-drift defined by
    \begin{gather}
        \vect{v}_{\nabla B} = \frac{m v^2_{\perp}}{2 q}\frac{\vect{B}\times \nabla B}{B^3}~.
        \label{eq:gradB}
    \end{gather}
    The $\nabla B$-drift varies thereby on scales which are larger compared to the Larmor radius. The direction of the $\nabla B$-drift depends on the species type.
    \item \textbf{Curvature Drift:}\\
    Due to centrifugal force acting on the particle in a curved magnetic filed the gyrocenter experience a curvature drift given by
    \begin{gather}
        \vect{v}_{\kappa} = \frac{m v^2_{\parallel}}{q}\frac{\vect{B}\times \vect{\kappa}}{B^2} = \frac{m v^2_{\parallel}}{q} \frac{\vect{B}\times \nabla B}{B^3}~;\qquad\vect{\kappa} = -(\vect{b}\cdot \nabla)\vect{b} = \frac{\nabla B}{B}~,
        \label{eq:curvature}
    \end{gather}
    where $\vect{b}$ is the unit vector along the magnetic field. To obtain the result for the curvature $\vect{\kappa}$ in Eq. (\ref{eq:curvature}) the plasma pressure has to be small in comparison die the magnetic field strength $B$. In the form of Eq. (\ref{eq:curvature}) $\nabla B$ and curvature drift can be treated similarly. \cite{Wesson2011}
\end{enumerate}

### 2023-05-09 bachelorthesis/22-confinement.tex
#39: Theory finished and closure corrected, First Draft finished

In a tokamak devices strong magnetic fields confine the hot plasma. As mentioned in Chapter \ref{sec:motion} a magnetic field forces a perpendicular particle motion and a motion containing gyro motion and slow perpendicular gyro center drifts. Because of the much smaller size of the Larmor radius compared to the device size $R_0$ the particle and energy losses are caused by the gyro center drift. To avoid this at the end of the field lines the magnetic field in the tokamak devices is shaped like a torus. This type of geometry has nested surfaces with constant magnetic flux, so-called \textit{flux sufaces} and with magnetic field lines which appearing on that surfaces.

>>>>

In a tokamak devices strong magnetic fields confine the hot plasma. As mentioned in Chapter \ref{sec:motion} a magnetic field forces a perpendicular particle motion and a motion containing gyro motion and slow perpendicular gyro center drifts. Because of the much smaller size of the Larmor radius compared to the device size $R_0$ the particle and energy losses are caused by the gyro center drift. To avoid this at the end of the field lines the magnetic field in the tokamak devices is shaped like a torus. This type of geometry has nested surfaces with constant magnetic flux, so-called \textit{flux sufaces} and with magnetic field lines which appearing on that surfaces. For stability reasons the magnetic field has a toroidal and a poloidal component which also holds equivalence of the plasma pressure. \cite{Stroth2011, Wesson2011} The toroidal component is produced by external coils whereas the poloidal component is provided by the toroidal plasma current. Together the components results in a magnetic field which follows helical trajectories [Fig \ref{fig:confinement}]. To characterize the the quality of confinement the so-called \textit{plasma beta} is used and is given by
\begin{gather}
    \beta = \frac{nT}{\mu_0 B^2/2}~.
\end{gather} 
with $n$ the plasma density, $T$ as temperature, $\mu_0$ the permeability in vacuum and the magnetic field strength $B$. Respectively, the plasma beta compares the thermal plasma pressure $nT$ to the ambient magnetic field pressure $\mu_0 B^2/2$ which indicates when the plasma beta is smaller the confinement is better. In a tokamak reactor the plasma beta has a typical order of few percent. \cite{Wesson2011}
\includegraphicsHere{Theory/Tokamak-Torus.pdf}{
	Toroidal flux surfaces in tokamak plasma with helical magnetic field (\textcolor{Ubtgreen}{green} line) in torus coordinates ($\rho$ (radial), $\phi$ (toroidal), $\theta$ (poloidal)) or zylinder coordinates ($Z$, $R$, $\phi$). \cite{Barton2015}
}{fig:confinement}{0.7}

### 2023-05-09 bachelorthesis/23-iontempgrad.tex
#39: Theory finished and closure corrected, First Draft finished

\newpage

\section{Ion Temperature Gradient (ITG) Driven Instability}
\label{sec:ITG}

The described electrostatic forces acting on the tokamak plasma causes a variety of collective phenomena. Microinstabilities is one of this phenomena which are instabilities on the spatial scale of the Larmor radius and frequencies much smaller than the cyclotron frequency. These instabilities are driven bei density and temperature gradients that are present in the tokamak devices. Turbulence caused from microinstabilities are considered as the main reason for the loss of particles energy and momentum from tokamak plasma Ref (10,11,12) which limits the confinement time of tokamak reactors. The most dominante instability in tokamak plasma is the ion temperature gradient (ITG) driven instability Ref (13,14,15).

\includegraphicsHere{Theory/ITG-Turbulence.png}{
	Sketch of ion temperature gradient driven instability on the outboard low field sid of the tokamak
}{fig:ITG}{0.6}

The ITG driven is present when a sufficiently large ion temperature gradient occurs and is driven by the so-called bad curvature of the outboard side of the tokamak. The mechanism will be briefly explained with the help of Fig. \ref{fig:ITG} Ref (18,17,16 and 4 in BA). An ion temperature perturbation ($T^\pm \gtrless 0$ \textcolor{blue}{blue} and \textcolor{red}{red} cells) on the outboard low field side of the tokamak causes the formation of curvature and $\nabla B$ dirft $\vect{v}_\mathrm{d}$ which compresses the plasma ($n_\pm \gtrless 0$, grey cells). This relates to the perturbation of the potential ($\phi^\pm  \gtrless 0$) in the adiabatic electron response. As stated in the equation for the electric field ($\vect{E} = -\nabla \phi$) the perturbation of the potential causes an electric field which results in a $\exb$ drift $\vect{v}_E$. The $\exb$-drifts direction transports hot plasma in the outer hot region of perturbation ($n^+ > 0$, \textcolor{red}{red} cell) and cooler plasma into inner cold region perturbation ($n^- < 0$, \textcolor{blue}{blue} cell) which reinforce the process and instability grows.

The described mechanism causes perturbation on the outboard side of the tokamak this region has a \enquote{bad curvature} and instability happens there. On the inboard side the opposing $\nabla B$ and $\nabla T$ results in a decay of perturbation which is referred to as \enquote{good curvature}. Furthermore such turbulent structures are called \textit{eddies} and the have a typical radial scale of serval Larmor radius. (Ref 19 BA). \bigskip

Through the explanation above is clear that the initial perturbation is growing when the normalized temperature gradient overcomes a critical value $R/L_\mathrm{T,c}$. For nonlinear turbulence the critical value is given by $R/L_\mathrm{T,c} \sim 6$. (Ref 3, 7 BA)

>>>>

\newpage
\section{Ion Temperature Gradient (ITG) Driven Instability}
\label{sec:ITG}

The described electrostatic forces acting on the tokamak plasma causes a variety of collective phenomena. Microinstabilities is one of this phenomena which are instabilities on the spatial scale of the Larmor radius and frequencies much smaller than the cyclotron frequency. These instabilities are driven bei density and temperature gradients that are present in the tokamak devices. Turbulence caused from microinstabilities are considered as the main reason for the loss of particles energy and momentum from tokamak plasma \cite{Brizard2007, Garbet2010, Horton1999} which limits the confinement time of tokamak reactors. The most dominante instability in tokamak plasma is the ion temperature gradient (ITG) driven instability \cite{Coppi1967, Cowley1991, Rudakov1961}

\includegraphicsHere{Theory/ITG-Turbulence.pdf}{
	Sketch of ion temperature gradient driven instability on the outboard low field sid of the tokamak \cite{Casson_PHD}
}{fig:ITG}{0.5}

The ITG driven is present when a sufficiently large ion temperature gradient occurs and is driven by the so-called bad curvature of the outboard side of the tokamak. The mechanism will be briefly explained with the help of Fig. \ref{fig:ITG} \cite{Beer_PHD, Casson_PHD, Dannert_PHD}. An ion temperature perturbation ($T^\pm \gtrless 0$ \textcolor{blue}{blue} and \textcolor{red}{red} cells) on the outboard low field side of the tokamak causes the formation of curvature and $\nabla B$ dirft $\vect{v}_\mathrm{d}$ which compresses the plasma ($n_\pm \gtrless 0$, \textcolor{gray}{gray} cells). This relates to the perturbation of the potential ($\phi^\pm  \gtrless 0$) in the adiabatic electron response. As stated in the equation for the electric field ($\vect{E} = -\nabla \phi$) the perturbation of the potential causes an electric field which results in a $\exb$ drift $\vect{v}_E$. The $\exb$ drifts direction transports hot plasma in the outer hot region of perturbation ($n^+ > 0$, \textcolor{red}{red} cell) and cooler plasma into inner cold region perturbation ($n^- < 0$, \textcolor{blue}{blue} cell) which reinforce the process and instability grows.\bigskip

The described mechanism causes perturbation on the outboard side of the tokamak this region has a \enquote{bad curvature} and instability happens there. On the inboard side the opposing $\nabla B$ and $\nabla T$ results in a decay of perturbation which is referred to as \enquote{good curvature}. Furthermore such turbulent structures are called \textit{eddies} and the have a typical radial scale of serval Larmor radius. \cite{Newins2006}. 
Through the explanation above is clear that the initial perturbation is growing when the normalized temperature gradient overcomes a critical value $R/L_\mathrm{T,c}$. For nonlinear turbulence the critical value is given by $R/L_\mathrm{T,c} \sim 6$. \cite{Dimits2000, Isliker2010}

### 2023-05-09 bachelorthesis/24-zonalflows.tex
#39: Theory finished and closure corrected, First Draft finished

Zonal flows are referred in the tokamak plasma to the $\exb$-flows [Eq. \ref{eq:exb}] that is connected to the on flux-surfaces constant electrostatic potential, also called \textit{zonal potential}. (Ref 28) Because of the variation only across flux-surfaces only the zonal flow is tangential to the flux-surfaces and do not contribute to the turbulent transport across flux-surfaces caused by the $\exb$-motion. 
Zonal flows are driven nonlinear through turbulent Reynolds stresses (Ref 32). For more about the mechanism behind zonal flow generation the reader is referred to Ref. (42,43,44,10). \bigskip

Furthermore zonal flows are connected with the radially alternating zonal potential what introduces an velocity shear,i.e., a radial variation of the zonal flow velocity. This causes a shear deformation of turbulent structures which is considered to play a significant role in the suppression of turbulence (Ref 29, 67) especially the ion temperature gradient driven turbulence (Ref 22, 63, 64, 65, 66). The nonlinear generation mechanism causes an shift, also known as Dimits shift (Ref 67), of the critical temperature gradient $R/L_\mathrm{T,c}$ which leads to an later development of ITG driven turbulence. The shearing process tilted und stretches turbulent eddies, caused be the shear flow, until they break into smaller eddies with reduced radial correlation length which characterize the shearing as decorrelation process. Turbulence will then be reduced because of this process. (Ref 29,68,28)  \bigskip

When the zonal flows are strong enough to suppress turbulence completely due to the shearing process, pattern formation occurs in the stabilized tokamak plasma the so-called $\exb$ staircase structure. The $\exb$ staircase pattern has following properties:
\begin{enumerate}
	\item[(1)] They occur on a \textit{radial mesoscale} of order of 10 \rhoth which is larger than the Larmor radius but smaller than the machine size $R_0$.
	\item[(2)] The structure are \textit{quasi stationary} in space and varies on time scales much longer than the typical turbulent time scales.
	\item[(3)] They have an \textit{typical amplitude} in terms of the $\exb$-shearing rate of the order of $10^{-1} \vth/R_0$.
	\item[(4)] Turbulent transport occurs in \textit{avalanches} with propagation strongly linked to the local $\exb$ shearing rate. (Ref 23)
\end{enumerate} \bigskip

The $\exb$ staircase pattern is manifest as radial structure formation in the $\exb$ shearing rate defined by\cite{rath2016, doi:10.1063/1.3005380, peeters2016}
\begin{equation}
	\wexb = \frac{1}{2} \frac{\partial^2 \langle \phi \rangle}{\partial \xcoord^2},
	\label{eq:shearingrate}
\end{equation}
where $\langle \phi \rangle$ is the zonal electrostatic potential normalized by $\rho_\ast T/e$ ($\rho_\ast = \rhoth/R$ is the thermal Larmor radius normalized with the major radius $R$, $T$ is the temperature, $e$ is the elementary charge).
The zonal potential is calculated from the electrostatic potential $\phi$ on the two-dimensional $\xcoord$-$\ycoord$-plane at the low field side according to\cite{rath2021}
\begin{equation}
\langle \phi \rangle = \frac{1}{L_\ycoord} \int_0^{L_\ycoord} \mathrm{d}\ycoord ~ \phi(\xcoord,\ycoord,s=0).
\end{equation}
The $\exb$ shearing rate $\wexb$ is the radial derivative of the advecting zonal flow velocity \cite{doi:10.1063/1.871313, doi:10.1063/1.872847} and quantifies the zonal flow induced shearing of turbulent structures \cite{doi:10.1063/1.859529, doi:10.1063/1.871313, doi:10.1063/1.872367}. \\ \bigskip

As mentioned before the turbulent transport occurs through avalanches which also mediate the turbulence of the ITG driven turbulence in the adabatic electron approximation. Avalanches are initial when the$\exb$ sharing rate has a zero crossing with a steep flank and propagates outwards with negative shearing and inwards for positive shearing (Ref 26,27). The fully developed avalanche resluts in almost completely quenched turbulence. Its development results in a discontinuous step int he dependency of turbulent fluxes on the temperature gradient als known as \textit{finite heat flux threshold} (Ref 25,77)

>>>>

Zonal flows are referred in the tokamak plasma to the $\exb$ flows [Eq.( \ref{eq:exb})] that is connected to the on flux-surfaces constant electrostatic potential, also called \textit{zonal potential}. \cite{Diamond2005} Because of the variation only across flux-surfaces only the zonal flow is tangential to the flux-surfaces and do not contribute to the turbulent transport across flux-surfaces caused by the $\exb$ motion. 
Zonal flows are driven nonlinear through turbulent Reynolds stresses \cite{Diamond1991}. For more about the mechanism behind zonal flow generation the reader is referred to Ref. \citenum{Brizard2007, Dubin1983, Hasegawa1978, Krommes2000}. \bigskip

\includegraphicsHere{Theory/Zonal-Flow-Generation.pdf}{
	Subdued turbulence because if stretched eddies which breaks in smaller structures caused by sheared $\exb$ drifts \cite{Hammett2009}
}{fig:zonalflow}{0.9}

Furthermore, zonal flows are connected with the radially alternating zonal potential what introduces a velocity shear, i.e., a radial variation of the zonal flow velocity. This causes a shear deformation of turbulent structures which is considered to play a significant role in the suppression of turbulence \cite{Biglari1990, Dimits2000} especially the ion temperature gradient driven turbulence \cite{Nakata2012, Makwana2014, Maeyama2014, Whelan2018, Whelan2019}. The nonlinear generation mechanism causes a shift, also known as Dimits shift \cite{Dimits2000}, of the critical temperature gradient $R/L_\mathrm{T,c}$ which leads to a later development of ITG driven turbulence. The shearing process tilted und stretches turbulent eddies, caused be the shear flow, until they break into smaller eddies with reduced radial correlation length [Fig. \ref{fig:zonalflow}] which characterize the shearing as decorrelation process. Turbulence will then be reduced because of this process. \cite{Biglari1990,Diamond2005,Burnell1997} \bigskip

\newpage

When the zonal flows are strong enough to suppress turbulence completely due to the shearing process, pattern formation occurs in the stabilized tokamak plasma the so-called $\exb$ staircase structure. The $\exb$ staircase pattern has the following properties:
\begin{enumerate}
	\item[(1)] They occur on a \textit{radial mesoscale} of order of 10 \rhoth which is larger than the Larmor radius but smaller than the machine size $R_0$.
	\item[(2)] The structure are \textit{quasi stationary} in space and varies on time scales much longer than the typical turbulent time scales.
	\item[(3)] They have an \textit{typical amplitude} in terms of the $\exb$-shearing rate of the order of $10^{-1} \vth/R_0$.
	\item[(4)] Turbulent transport occurs in \textit{avalanches} with propagation strongly linked to the local $\exb$ shearing rate. \cite{McMillan2009}
\end{enumerate} \bigskip

The $\exb$ staircase pattern is manifest as radial structure formation in the $\exb$ shearing rate defined by\cite{Rath2016, Pueschel2008, Peeters2016}
\begin{equation}
	\wexb = \frac{1}{2} \frac{\partial^2 \langle \phi \rangle}{\partial \xcoord^2},
	\label{eq:shearingrate}
\end{equation}
where $\langle \phi \rangle$ is the zonal electrostatic potential normalized by $\rho_\ast T/e$ ($\rho_\ast = \rhoth/R$ is the thermal Larmor radius normalized with the major radius $R$, $T$ is the temperature, $e$ is the elementary charge).
The zonal potential is calculated from the electrostatic potential $\phi$ on the two-dimensional $\xcoord$-$\ycoord$-plane at the low field side according to\cite{Rath2021}
\begin{equation}
\langle \phi \rangle = \frac{1}{L_\ycoord} \int_0^{L_\ycoord} \mathrm{d}\ycoord ~ \phi(\xcoord,\ycoord,s=0).
\end{equation}
The $\exb$ shearing rate $\wexb$ is the radial derivative of the advecting zonal flow velocity \cite{Hahm1995, Waltz1998} and quantifies the zonal flow induced shearing of turbulent structures \cite{Biglari1990, Hahm1995, Burnell1997}. \\ \bigskip

As mentioned before the turbulent transport occurs through avalanches which also mediate the turbulence of the ITG driven turbulence in the adabatic electron approximation. Avalanches are initial when the $\exb$ sharing rate has a zero crossing with a steep flank and propagates outwards with negative shearing and inwards for positive shearing \cite{Idomura2009,McMillan2009}. The fully developed avalanche resluts in almost completely quenched turbulence. Its development results in a discontinuous step int he dependency of turbulent fluxes on the temperature gradient als known as \textit{finite heat flux threshold}. \cite{Peeters2016,Weikl2017}
